Towards Spherical Robots for Mobile Mapping in Human Made Environments\tnoteref{mytitlenote}

### 2021-07-12 Latex/rebuttal.txt

> 
> Reviewer 2: The authors have written a very interesting paper on the
> topic of mobile mapping with a spherical robot in human-made
> environments. In a rather slim paper and pleasantly concise form,
> they present their approach and show first results based on
> simulated and real measurement data. The reading flow is mostly
> pleasant and no irrelevant information is given. Conversely, the
> authors leave out important details in several places, which would
> be very helpful for the reader's understanding. In the following,
> there are several points for improvement and critical questions that
> should be taken into account by the authors. Due to the large number
> and the importance of central points, I recommend a major revision.
> 
> - A short outline at the end of the introduction is missing and
> would give the reader a better overview of the further content and
> should therefore be added.

We added a short outline, introducing you to the next sections.

> - In the entire article, no reference is made to the targeted or
> required accuracy potential with regard to the quality of the pose
> determination and the final point cloud, which in my view requires
> at least a rough classification.

Unfortunatley we could not extract a 'ground trouth' scan for the real world datasets, due to covid pandemic issues. Both rooms were inaccessible, thus we evaluate the accuracy of the real world dataset on a later point in time. 
The accuracy potential depends heavily on the motion profile, the environment
and the hardware being used, thus is not easy to generalize. However, the
simulation results show that the algorithm has the potential to improve the map
quality based on point-distances by approximately a factor of ten. We added this
to the conclusions.

### 2021-07-13 Latex/sections/registration.tex
equation -> align

The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to infinitely extending, $k$-th plane in 3D vector space
\begin{equation}\label{eq:hesse}
	D_{h}^{i,k} = \vec{n}_{\mathcal{P}_k} \cdot \left[ T(\vec{p}_i) - \vec{a}_{\mathcal{P}_k} \right] ,
\end{equation}
with normal $\vec{n}_{\mathcal{P}_k}$ and supporting point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
The distance $D_{h}^{i,k}$ of the $i$-th point to the $k$-th plane reflects the length of the line segment, constructed from the $i$-th point to its projection on the $k$-th plane.
Therefore, by definition the line segment is parallel to the normal vector of the plane.
Thus, the $i$-th points projection $\widetilde{T}(\vec{p}_i)$ onto the $k$-th plane, which is required later, is easily calculated by shifting the point against normal direction:
\begin{equation}\label{eq:projection}
	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \vec{n}_{\mathcal{P}_k}
\end{equation}

The simplicity of the Hesse distance is at contrast with its inability to take the extent of the plane into account.
Therefore, a second distance model is introduced: the polygon projection distance (PPD).
The convex hull represents the extend of a plane by forming a polygon with the outer most points that are assigned to a given plane. 
It is able to represent the expansion of the plane in all directions, and thus is utilized as a distance model.
To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e. $\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in~\ref{fig:proj}).
Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the maximal component of the normal vector of the plane.
E.g. in figure~\ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
Sunday et al.~\cite{pga} presents an optimal algorithm to check whether the point projection lies inside the polygon in 2D. 
If the point is inside the polygon, its PPD is set to zero.
If, however, the point is outside the polygon, the shortest distance to the polygon is calculated by looking at the minimum distance of the $i$-th point to each line segment, making up the convex hull of the $k$-th plane in 3D.
Let the polygon $P_k$ be made up of line segments $s_{m, j}$.
The line segment $s_{m, j}$ consists of the points $p_m$ and $p_j$.
The line segment is parameterized by
\begin{equation}
\label{eq:line}
	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
\end{equation}
where $t \in \left[ 0, 1 \right]$. 
We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
\begin{equation}
	\label{eq:dist}
	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
\end{equation} 
It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
Therefore, we set
\begin{equation}
	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
\end{equation}  
resulting in 
\begin{equation}
	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
\end{equation}
Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
In that case, the projection of the point $p_i$ onto the line given by eq.~\ref{eq:line} is not between $p_j$ and $p_m$. 
Instead, its projection onto the line falls outside of the segment.
By constraining $t_0$ we get 
\begin{equation}
\label{eq:minmax}
	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
\end{equation}
which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.

@@

The presented algorithm employs LPC on each individual line scan.
The idea is to find planar features utilizing an optimized approximate k nearest neighbor (AKNN) search for normal calculation as in~\cite{aknnbbs}.
Thus, normals are calculated for every point in the respective line scan.
Then, the local normal model is combined with a region growing approach.
The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
A region grows if one of these neighbors has a distance shorter then a threshold $d_{growth}$ and additionally has a similar normal.
We define similarity of two normals via the angle between them, which is:
\begin{equation}
\label{eq:smallestalpha}
	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
\end{equation}
However, the \textit{smallest} angle between two \textit{normal} vectors is 
\begin{equation}
\label{eq:alphahat}
\hat \alpha(\vec{n}_1, \vec{n}_2) = 
\begin{cases}
	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
\end{cases}
\; ,
\end{equation}
because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.

@@

Assuming we now know the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
The transformation $T(\vec{p}_i)$ of each point $\vec{p}_i$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}. Transforming the result back from homogeneous coordinates and using $C_a$ and $S_a$ to denote the cosine and sine of the angle in the subscript, and $t_a$ denoting the translation along the axis in the subscript yields:
\begin{align}
	T(\vec{p}_i)  =
    \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
        x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x\\
        x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y\\
        x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z
    \end{bmatrix}$}
\end{align}
From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_{i})$ that computes the distance of a point $\vec{p}_{i}$ to its corresponding plane $\mathcal{P}_k$.
Omitting the arguments of the function for simplicity:
\begin{equation}
\begin{split}
    D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
      &= n_{\mathcal{P}_k}^x (x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x)\\
       &+ n_{\mathcal{P}_k}^y(x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
       &\qquad+ y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y)\\
       &+ n_{\mathcal{P}_k}^z (x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
       &\qquad+ y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z)\\
       &- \rho_{\mathcal{P}_k}
\end{split}
\end{equation}

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{equation}
\begin{split}
\nabla E_\varphi = &n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]\\
			       & +y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
                   &+ n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\
                   &+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
\end{split} \;\; ,
\end{equation}
\begin{equation}
\begin{split}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
\end{split}
\;\; ,
\end{equation} 
\begin{equation}
\begin{split}
	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi]\\
       &+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
\end{split}
\end{equation}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.

>>>>

The Hesse-normal form describes the distance from the $i$-th transformed point $T(\vec{p}_i)$ to infinitely extending, $k$-th plane in 3D vector space
\begin{align}\label{eq:hesse}
	D_{h}^{i,k} = \vec{n}_{\mathcal{P}_k} \cdot \left[ T(\vec{p}_i) - \vec{a}_{\mathcal{P}_k} \right] ,
\end{align}
with normal $\vec{n}_{\mathcal{P}_k}$ and supporting point $\vec{a}_{\mathcal{P}_k}$ of the $k$-th plane.
The distance $D_{h}^{i,k}$ of the $i$-th point to the $k$-th plane reflects the length of the line segment, constructed from the $i$-th point to its projection on the $k$-th plane.
Therefore, by definition the line segment is parallel to the normal vector of the plane.
Thus, the $i$-th points projection $\widetilde{T}(\vec{p}_i)$ onto the $k$-th plane, which is required later, is easily calculated by shifting the point against normal direction:
\begin{align}\label{eq:projection}
	\widetilde{T}(\vec{p}_i) = T(\vec{p}_i) - D_{h}^{i,k} \vec{n}_{\mathcal{P}_k}
\end{align}

The simplicity of the Hesse distance is at contrast with its inability to take the extent of the plane into account.
Therefore, a second distance model is introduced: the polygon projection distance (PPD).
The convex hull represents the extend of a plane by forming a polygon with the outer most points that are assigned to a given plane. 
It is able to represent the expansion of the plane in all directions, and thus is utilized as a distance model.
To find the PPD, first a corresponding point gets projected onto the infinitely extending plane representation given by the Hesse form, i.e. $\widetilde{T}(\vec{p}_k)$ is found from eq.~\eqref{eq:projection} (green point in~\ref{fig:proj}).
Then, the 3D polygon, i.e. the points that make up the convex hull, as well as the corresponding 3D point projection, are projected again into a 2D vector space, using the maximal component of the normal vector of the plane.
E.g. in figure~\ref{fig:proj}, the blue planes normal vector has its mayor component in z-direction (upwards), thus the projection polygon distance is calculated in the xy-plane.
Using the maximum of the absolute magnitude of the normal vector ensures that the most sensible 2D projection is used for every direction.
Sunday et al.~\cite{pga} presents an optimal algorithm to check whether the point projection lies inside the polygon in 2D. 
If the point is inside the polygon, its PPD is set to zero.
If, however, the point is outside the polygon, the shortest distance to the polygon is calculated by looking at the minimum distance of the $i$-th point to each line segment, making up the convex hull of the $k$-th plane in 3D.
Let the polygon $P_k$ be made up of line segments $s_{m, j}$.
The line segment $s_{m, j}$ consists of the points $p_m$ and $p_j$.
The line segment is parameterized by
\begin{align}
\label{eq:line}
	s_{m, j}(t) = p_m + t \cdot \left( p_j - p_m \right) ,  
\end{align}
where $t \in \left[ 0, 1 \right]$. 
We set up a distance function, which measures the distance from the query point $p_i$ to an arbitrary point on the line segment.
\begin{align}
	\label{eq:dist}
	d_{m, j}(t) = \lVert s_{m, j}(t) - p_i \rVert 
\end{align} 
It is now possible to find the shortest distance to the line segment by finding the argument of the function that minimizes this distance.
Therefore, we set
\begin{align}
	\pardiff{t} d_{m, j}(t)^2 \overset{!}{=} 0 ,
\end{align}  
resulting in 
\begin{align}
	t_0 = \frac{\left( p_j - p_m \right) \cdot \left( p_i - p_m \right)}{\left( p_j - p_m \right)^2} .
\end{align}
Note the possibility of $t_0 \notin \left[ 0, 1 \right]$. 
In that case, the projection of the point $p_i$ onto the line given by eq.~\ref{eq:line} is not between $p_j$ and $p_m$. 
Instead, its projection onto the line falls outside of the segment.
By constraining $t_0$ we get 
\begin{align}
\label{eq:minmax}
	\hat t = \min \left( \max \left( t_0, 0\right) , 1 \right) ,
\end{align}
which is the argument that gives the shortest distance to the line segment, when inserted to eq. \ref{eq:dist}.
We find the PPD from the point $p_i$ to the polygon $P_k$ by calculating the minimum shortest distance over all line segments that make up the polygon.

@@

The presented algorithm employs LPC on each individual line scan.
The idea is to find planar features utilizing an optimized approximate k nearest neighbor (AKNN) search for normal calculation as in~\cite{aknnbbs}.
Thus, normals are calculated for every point in the respective line scan.
Then, the local normal model is combined with a region growing approach.
The region growing is implemented by storing the k-nearest neighbors in a queue, which represents the iteration order of the clustering algorithm.
A region grows if one of these neighbors has a distance shorter then a threshold $d_{growth}$ and additionally has a similar normal.
We define similarity of two normals via the angle between them, which is:
\begin{align}
\label{eq:smallestalpha}
	\alpha(\vec{n}_1, \vec{n}_2) = \arccos(\vec{n}_1 \cdot \vec{n}_2) \; .
\end{align}
However, the \textit{smallest} angle between two \textit{normal} vectors is 
\begin{align}
\label{eq:alphahat}
\hat \alpha(\vec{n}_1, \vec{n}_2) = 
\begin{cases}
	2\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{3}{2} \pi \\
	\alpha(\vec{n}_1, \vec{n}_2) - \pi & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \pi \\
	\pi - \alpha(\vec{n}_1, \vec{n}_2) & \text{if  } \alpha(\vec{n}_1, \vec{n}_2) > \frac{\pi}{2} 
\end{cases}
\; ,
\end{align}
because the opposing normals $-\vec{n}_1$ and $-\vec{n}_2$ always have to be considered, since they correspond to the same plane.
We say that two normals are similar if the angle $\hat \alpha$ from eq. \ref{eq:alphahat} is smaller than a threshold $\epsilon_{\alpha}$.

@@

Assuming we now know the Hesse-normal form of all planes and all assigned points to these planes, we register the points to have a minimal distance to their respective planes.
The transformation $T(\vec{p}_i)$ of each point $\vec{p}_i$ with respect to a 6 DoF motion is described in homogeneous coordinates using the roll-pitch-yaw ($\varphi-\vartheta-\psi$) Tait-Brian angles as in~\cite{diebel2006representing}. Transforming the result back from homogeneous coordinates and using $C_a$ and $S_a$ to denote the cosine and sine of the angle in the subscript, and $t_a$ denoting the translation along the axis in the subscript yields:
\begin{align}
	T(\vec{p}_i)  =
    \resizebox{0.35\textwidth}{!}{$\begin{bmatrix}
        x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x\\
        x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta) + y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y\\
        x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta) + y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z
    \end{bmatrix}$}
\end{align}
From this we define the function $D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_{i})$ that computes the distance of a point $\vec{p}_{i}$ to its corresponding plane $\mathcal{P}_k$.
Omitting the arguments of the function for simplicity:
\begin{align}
\begin{split}
    D &= T(\vec{p}_i) \cdot \vec{n}_{\mathcal{P}_k} \\
      &= n_{\mathcal{P}_k}^x (x_i C_\vartheta C_\psi - y_iC_\vartheta S_\psi + z_i S_\vartheta + t_x)\\
       &+ n_{\mathcal{P}_k}^y(x_i (C_\varphi S_\psi + C_\psi S_\varphi S_\vartheta)\\
       &\qquad+ y_i(C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi) - z_i C_\vartheta S_\varphi + t_y)\\
       &+ n_{\mathcal{P}_k}^z (x_i(S_\varphi S_\psi - C_\varphi C_\psi S_\vartheta)\\
       &\qquad+ y_i(C_\psi S_\varphi + C_\varphi S_\vartheta S_\psi) + z_i C_\varphi C_\vartheta + t_z)\\
       &- \rho_{\mathcal{P}_k}
\end{split}
\end{align}

This distance function is what we want to minimize for all points and their respective planes.
Hence the error function $E$ is chosen as the square of the L2-norm of the distance:
\begin{align}
	E = \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \norm[2]{D(\varphi,\vartheta,\psi,t_x,t_y,t_z, \vec{p}_i)}^2
\end{align}
Its gradient follows then immediately:
\begin{align}
	&\nabla E =  \sum_{\forall \mathcal{P}_k}\sum_{\vec{p}_i \in \mathcal{P}_k} \resizebox{0.28\textwidth}{!}{$\begin{bmatrix}\pardiff{\varphi} E&\pardiff{\vartheta} E&\pardiff{\psi} E&\pardiff{t_x} E&\pardiff{t_y} E&\pardiff{t_z} E\end{bmatrix}^T$}\\
    &\Rightarrow\nabla E = \resizebox{0.35\textwidth}{!}{$\sum_{\forall \mathcal{P}_k} \sum_{\vec{p}_i\in \mathcal{P}_k}2D(\vec{\Pi}, \vec{p}_i)\begin{bmatrix}\nabla E_\varphi &\nabla E_\vartheta & \nabla E_\psi & n_{\mathcal{P}_k}^x&n_{\mathcal{P}_k}^y&n_{\mathcal{P}_k}^z\end{bmatrix}^T$}
\end{align}
Where
\begin{align}
\begin{split}
\nabla E_\varphi = &n_{\mathcal{P}_k}^y(x_i[-S_\varphi S_\psi+C_\varphi C_\psi S_\vartheta]\\
			       & +y_i[-S_\varphi C_\psi -C_\varphi S_\vartheta S\psi] - z_i C_\varphi C_\vartheta)\\
                   &+ n_{\mathcal{P}_k}^z(x_i[C_\varphi S_\psi+C_\varphi C_\psi S_\vartheta] \\
                   &+y_i[C_\varphi C_\psi -S_\varphi S_\vartheta S\psi] - z_i S_\varphi C_\vartheta)
\end{split} \;\; ,
\end{align}
\begin{align}
\begin{split}
 \nabla E_\vartheta = & n_{\mathcal{P}_k}^x(-x_iS_\vartheta C_\psi + y_i S_\vartheta S_\psi + z_iC_\vartheta) \\
        &+ n_{\mathcal{P}_k}^y(x_iC_\psi S_\varphi C_\vartheta - y_iS_\varphi C_\vartheta S_\psi + z_i S_\vartheta S_\varphi)  \\
        &+ n_{\mathcal{P}_i}^z(-x_iC_\varphi C_\psi C_\vartheta + y_iC_\varphi C_\vartheta S_\psi - z_i C_\varphi 
\end{split}
\;\; ,
\end{align} 
\begin{align}
\begin{split}
	\nabla E_\psi = &n_{\mathcal{P}_k}^x\left(-x_iC_\vartheta S_\psi - y_iC_\vartheta C_\psi\right) \\
       &+ n_{\mathcal{P}_k}^y(x_i[C_\varphi C_\psi - S_\varphi S_\vartheta S_\psi]\\
       &+ y_i[-C_\varphi S_\psi - S_\varphi S_\theta C_\psi]) \\
       &+ n_{\mathcal{P}_k}^z(x_i[S_\varphi C_\psi + C_\varphi S_\psi S_\theta] \\
       &+ y_i[-S_\psi S_ \varphi + C_\varphi S_\vartheta C_\psi]) 
\end{split}
\end{align}
and $\vec{\Pi}=\begin{bmatrix}\varphi & \vartheta & \psi & t_x & t_y & t_z\end{bmatrix}^T$.
